---
title: "Covid Surival Rates based on GDP"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
```

## Import global Covid Cases and Deaths

```{r import_data, echo=TRUE}
root_url = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"

confirmed_global = read_csv(paste(root_url, "time_series_covid19_confirmed_global.csv", sep=""))
deaths_global = read_csv(paste(root_url, "time_series_covid19_deaths_global.csv", sep=""))
```

## Tidy Data

-   Convert from date columns to row columns

-   Remove unnecessary columns and rename columns

-   Merge cases and deaths

-   Add population column

-   Convert data character to date format

-   Limit the year to just 2020

-   Group by Country/Region

```{r combine_global_confirmed_and_death, echo=TRUE}
deaths_global <- deaths_global %>%
    pivot_longer(cols = -c(`Province/State`,
                           `Country/Region`, Lat, Long),
                names_to = "date",
                values_to = "deaths") %>%
    select(-c(Lat, Long))

confirmed_global <- confirmed_global %>%
    pivot_longer(cols = -c(`Province/State`,
                           `Country/Region`, Lat, Long),
                names_to = "date",
                values_to = "cases") %>%
    select(-c(Lat, Long))

global <- confirmed_global %>%
    full_join(deaths_global) %>%
    rename(Country_Region = `Country/Region`,
           Province_State = `Province/State`) %>%
    mutate(date = mdy(date))

global = global[as.numeric(format(global$date, "%Y")) == 2020, ]
global <- global %>%
  group_by(Country_Region) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths))

uid = read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv") %>%
  select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))

uid <- uid %>%
  group_by(Country_Region) %>%
  summarise(Population = sum(Population))

global <- global %>%
    left_join(uid, by = c("Country_Region")) %>%
    select(Country_Region, cases, deaths, Population)


```

## Add  GDP Data

-   Import GDP Data

-   Remove unnecessary columns from GDP and rename columns

-   Merge with global

-   Add calculated columns

    -   GDP per Capita

    -   Cases per Capita

    -   Deaths per Capita

```{r per_capita, echo=TRUE}
gdp = read_csv("https://raw.githubusercontent.com/datasets/gdp/refs/heads/main/data/gdp.csv")
gdp <- gdp[gdp$Year == "2020", ]
gdp <- gdp %>%
  select(-c(Year, `Country Code`)) %>%
  rename(Country_Region = `Country Name`,
         GDP = Value)

global <- global %>%
    left_join(gdp, by = "Country_Region")


global$gdp_pop = global$GDP / global$Population
global$cases_pop = global$cases / global$Population
global$deaths_pop = global$deaths / global$Population
```

# Cases Per GDP (plot)

```{r plot_cases, echo=TRUE}
ggplot(global, aes(x = gdp_pop, y = cases_pop)) +
geom_point(size = 3) +
labs(x = "GDP per capita", y = "Cases per capita", title = "COVID-19 Cases vs. GDP Per Capita") +
theme_bw()

ggplot(global, aes(x = log10(gdp_pop), y = cases_pop)) +
geom_point(size = 3) +
labs(x = "Log10 of GDP per capita", y = "Cases per capita", title = "COVID-19 Cases vs. GDP Per Capita") +
theme_bw()
```

# Deaths Per GDP (plot)

```{r plot_deaths, echo=TRUE}
ggplot(global, aes(x = gdp_pop, y = deaths_pop)) +
geom_point(size = 3) +
labs(x = "GDP per capita", y = "Deaths per capita", title = "COVID-19 Deaths vs. GDP Per Capita") +
theme_bw()

ggplot(global, aes(x = log10(gdp_pop), y = deaths_pop)) +
geom_point(size = 3) +
labs(x = "Log10 of GDP per capita", y = "Deaths per capita", title = "COVID-19 Deaths vs. GDP Per Capita") +
theme_bw()
```

# GDP and Death Correlation?

## There might be a connection between deaths and gdp, lets compare deaths to cases

```{r death_per_case, echo=TRUE}

global$case_death_deaths_pop = global$deaths_pop / global$cases_pop

ggplot(global, aes(x = gdp_pop, y = case_death_deaths_pop)) +
geom_point(size = 3) +
labs(x = "GDP per capita", y = "Deaths/Cases per capita", title = "COVID-19 Deaths/Cases vs. GDP Per Capita") +
theme_bw()

ggplot(global, aes(x = log10(gdp_pop), y = case_death_deaths_pop)) +
geom_point(size = 3) +
labs(x = "Log10 of GDP per capita", y = "Deaths/Cases per capita", title = "COVID-19 Deaths/Cases vs. GDP Per Capita") +
theme_bw()

```

```{r trendline, echo=FALSE}
global <- global[complete.cases(global), ]
mod <- lm(gdp_pop ~ case_death_deaths_pop, data=global)
summary(mod)
global_mod <- global %>% mutate(pred = predict(mod))

global_mod %>% ggplot() +
  geom_point(aes(x = case_death_deaths_pop, y = gdp_pop), color ="blue") +
  geom_point(aes(x = case_death_deaths_pop, y = pred), color ="red")
```

# Conclusion

After looking at the data, we can see that there is not enough correlation for us to say that GDP and covid survival rate are related. After a close inspection of the data, there is no relationship. A relationship may be discovered if more variables are considered, but without any correlation whatsoever, it seems unlikely
