---
title: "NYPD Shootings as Seasonal Data"
date: "2024-11-19"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(forecast)
```

## NYPD Seasonal Shootings

### Import NYPD Shooting Data

```{r import_data, echo=TRUE}
url <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"

nyc_crimes <- read_csv(url)
```

## Tidy Data

### Convert Date from a character to a date

```{r convert_date, echo=TRUE}
nyc_crimes <- nyc_crimes %>% 
  mutate(OCCUR_DATE = as.Date(OCCUR_DATE, format = "%m/%d/%Y"))

```

## Visualize

### Plot shootings over time

```{r plot_by_month_year, echo=TRUE}
library(ggplot2)

ggplot(nyc_crimes, aes(x = OCCUR_DATE)) + 
  geom_histogram(binwidth = 1) + 
  labs(title = "NYPD Crimes Occurrences in 2018", 
       x = "Occurrence Date", 
       y = "Occurrences") + 
  theme_bw()
```

### Group by Month and Year then Plot

```{r plot_by_month_all_time, echo=TRUE}
annual_crime_count <- nyc_crimes %>% 
  group_by(month = floor_date(OCCUR_DATE, "month")) %>% 
  summarize(count = n())

ggplot(annual_crime_count, aes(x = month, y = count)) +
  geom_line() +
  labs(title = "NYC Crime Counts by month", x = "month", y = "Count") +
  theme_bw()
```

### Group by Month and Plot

```{r plot_by_month, echo=TRUE}
nyc_crimes %>%
    group_by(month = as.factor(month(nyc_crimes$OCCUR_DATE ))) %>%
    summarise(total_shootings = n()) %>%
    ggplot(aes(x = factor(month), y = total_shootings)) +
    geom_bar(stat = "identity") +
    labs(title = "Total Shootings by Month",
         x = "Month",
         y = "Total Shootings")
```

## Model

### Match a sine curve to the data

```{r sine_curve, echo=TRUE}
occurrences <- nyc_crimes %>% 
  group_by(month = month(OCCUR_DATE, label = TRUE),
           year = year(OCCUR_DATE)) %>% 
  summarize(n = n())

model <- lm(n ~ sin(2 * pi * n), data = occurrences)

summary(model)

ggplot(occurrences, aes(x = month, y = n)) +
  geom_line(aes(y = predict(model)), color = "red") +
  geom_point()
```


# Conclusion

#### The data matches the sine curve reasonably because it shows a very clear cyclic pattern of crime incidents throughout the year. It is also important to note that the data is not as comprehensive as it could be, as there are a variety of crimes that are not included in this data set. Nevertheless, it is still a good approximation of the overall trend of crime incidents in New York City.
